---
title: "Executive Summary of GM Findings"
author: Andrew Nalundasan
date: August 12, 2021
output: slidy_presentation
---

```{r echo = FALSE}
# Course: OMSBA 5210
# Purpose: Quick Project # 2 - Slides
# Date: August 12, 2021
# Author: Andrew Nalundasan
```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# NOTE, DO NOT CLEAR PACKAGES, IF YOU DO IT WILL NOT RUN
```


```{r message=FALSE, include=FALSE}

library(ggannotate)    # annotation tool
library(scales)        # to get $ and , on axis in graph
library(tidyverse)     # contains ggplot2, dplyr, and several other packages
library(vtable)        # NHK special
library(gghighlight)   # visual formatting - highlights
library(ggiraph)       # ggiraph
```


```{r include=FALSE}
# load and tidy data
# read data into environment and assign to variable
cereal_product <- read_csv("../02_raw_data/mtp_product_data.csv")
cereal_sales <- read_csv("../02_raw_data/mtp_sales_data.csv")

# create new column for 'company' for each cereal company
cereal_product <- cereal_product %>% 
  mutate(company = case_when(
    str_starts(brand, 'GENERAL') ~ 'General Mills', 
    str_starts(brand, 'KELLOGGS') ~ 'Kelloggs', 
    str_starts(brand, 'POST') ~ 'Post'))

# remove cereal company name from 'brand' column
cereal_product <- cereal_product %>% 
  mutate(brand = str_replace_all(brand, "GENERAL MILLS", "")) %>% 
  mutate(brand = str_replace_all(brand, "KELLOGGS", "")) %>% 
  mutate(brand = str_replace_all(brand, "POST", "")) %>% 
  mutate(flavor = stringr::str_to_title(flavor)) %>% 
  mutate(brand = str_to_title(brand)) %>% 
  mutate(brand = str_replace_all(brand, "Cinnamon Tst Cr", "Cinnamon Toast Crunch"))

# update UPC formatting  
cereal_product <- cereal_product %>%
  mutate(UPC = str_sub(UPC, 4, -1))

# update UPC formatting
cereal_product <- cereal_product %>% 
  mutate(UPC = str_replace_all(UPC, "-", "."))

# join cereal_product and cereal_sales together
cereal_ps <- left_join(cereal_product, cereal_sales, by = "UPC")

# data for graphs
cb_palette <- c("#999999", "#66CCCC", "#6600FF")

cereal_numeric <- cereal_ps %>% 
  mutate(company1 = case_when(
    str_starts(company, 'General') ~ "GM", 
    str_starts(company, 'Kelloggs') ~ "Kelloggs", 
    str_starts(company, 'Post') ~ "Post"))
```

Agenda
===============================================

1. Introduction

2. Finding 1

3. Finding 2

4. Key Takeaways

```{r eval=FALSE, include=FALSE}
# Hello everyone. Thank you for this opportunity to present to you the findings from my recent analysis for General Mills and our competitors. In this video, I will give a brief overview of the dataset and the analyses performed, I will speak about the findings and recommendations that came out of the analyses, and I'll wrap it up with our key takeaways. 
```



Introduction
===============================================

+ talk about the data here

+ talk about the analyses completed

+ pull stuff from memo



General Mills annual sales trending downward
===============================================

<div style="float: left; width: 35%;">

+ Market research shows that Kellogg's is our main competitor

+ Kellogg's currently the only cereal company with positive sales trends

+ What can General Mills do to stay competitive?

</div>

<div style="float: left; width: 65%;">

```{r echo=FALSE, message=FALSE, warning=FALSE}
# touch up scatter plot and save for memo

cb_palette <- c("#6600FF", "#66CCCC", "#999999")

company_sales_scatter <- cereal_ps %>% 
  group_by(week, company) %>% 
  summarise(count_n = n()) %>% 
  mutate(toolt = paste0('Company: ', company, 
                        '\nCount: ', count_n)) %>%
  ggplot(mapping = aes(x = week, y = count_n, color = company)) +
  scale_color_manual(values = cb_palette) +
  geom_point_interactive(aes(tooltip = toolt), alpha = 0.5) + 
  labs(x = "Week of the Year",
       y = "Number of Sales",
       title = "Annual sales trends for all cereal companies",
       subtitle = "Strategic ad campaigns could help influence positive sales trends") + 
  guides(color = "none", labels = "none") + 
  theme_classic() +
  theme(axis.ticks.x = element_blank()) + 
  theme(axis.ticks.y = element_blank()) +
  geom_smooth(method = "lm", se = FALSE) + 
  
  geom_text(data = data.frame(x = 48.5822707072111, y = 69.4390740593849, label = "Post"),
            mapping = aes(x = x, y = y, label = label),
            hjust = 0L, colour = "#999999", fontface = 2, inherit.aes = FALSE) + 
  
  geom_text(data = data.frame(x = c(47.7472046672058, 46.1814558421958 ),
                              y = c(282.469108673468, 167.977195797176),
                              label = c("Kellogg's", "General Mills")),
            mapping = aes(x = x, y = y, label = label),
            hjust = 0.2, colour = c("#66CCCC", "#6600FF"), fontface = 2, inherit.aes = FALSE)

# view the viz
ggiraph(ggobj = company_sales_scatter)

# make annotations
# ggannotate(company_sales_scatter)

# save the viz
ggsave(filename = "company_sales_scatter.png", plot = company_sales_scatter)
```


</div>

Update to advertisement strategies will boost number of sales
===============================================

<div style="float: left; width: 35%;">

+ Small ad campaigns are more effective than medium ad campaigns

+ Allocate advertisement budget to small ad campaigns to cut costs from medium ad campaigns

+ this shows the effect that ads have on sales

</div>

<div style="float: left; width: 65%;">

```{r echo=FALSE, message=FALSE, warning=FALSE}

cb_palette <- c("#6600FF", "#66CCCC")

weekly_ad_sales_smooth <- cereal_numeric %>% 
  mutate(ad = as.factor(ad)) %>% 
  filter(company == 'General Mills') %>% 
  filter(ad != "NONE") %>% 
  group_by(week, ad) %>% 
  summarise(count_n = n()) %>% 
  ggplot(mapping = aes(x = week, y = count_n, color = ad, legend = "none")) +
  scale_color_manual(values = cb_palette) + 
  geom_smooth_interactive(se = FALSE) +
  guides(color = "none", labels = "none") +  
  theme_classic() + 
  labs(x = "Week of the Year", 
       y = "Number of Sales", 
       title = "General Mills annual trends of advertisement campaigns",
       subtitle = "Small ads outperform medium ads all year long") + 
  
  geom_text(data = data.frame(x = 39.0202225326685, y = 4.6481581052289, label = "Medium ads"),
            mapping = aes(x = x, y = y, label = label),
            hjust = -0.35, vjust = 1L, colour = "#66CCCC",  fontface = 2, inherit.aes = FALSE) + 
  
  geom_text(data = data.frame(x = 43.1127335651892, y = 10.6604358222577, label = "Small ads"),
            mapping = aes(x = x, y = y, label = label),
            hjust = 0.25,colour = "#6600FF", fontface = 2, inherit.aes = FALSE) +

  geom_rect(data = data.frame(xmin = 35.107224456232, xmax = 40.979094351119, ymin = 3.1227927337657, 
                              ymax = 11.062435478814),
            mapping = aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            size = 0L, fill = "orange", alpha = 0.2, inherit.aes = FALSE)


#ggannotate(weekly_ad_sales_smooth)

# view the viz
ggiraph(ggobj = weekly_ad_sales_smooth)
```


</div>


Key Takeaways
===============================================

+ Update advertisement strategies to focus on small campaigns over medium campaigns

+ Costs saved from medium campaigns can be focused elsewhere within the company

+ Verify increase in sales due to updated strategies

+ Drive GM sales trends to be positive, then focus energy on competition within the market
