---
title: "Week-8-Notes"
author: "Andrew Nalundasan"
date: "8/13/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview video

+ Forget about plotly

    - Use ggiraph for interactive graphs
        - example in week 7 overview video
    - Use dygraph for line graphs
    
+ Focus on Example 3 for Shiny Interactive Visual

    - Use 'global chunk' approach

+ Start thinking about DTC

# HTML Widgets

```{r}
# load libraries
library(tidyverse)
library(vtable)
library(dygraphs)
library(xts)
library(gganimate)
library(gapminder) # animation data
```


## What are they?

```{r}
# load data
qp1 <- read_csv("../02_raw_data/qp1_data.csv")

# convert variables
qp1$zip_code <- as.factor(qp1$zip_code)

vtable(qp1)

attach(qp1)
```

+ HTML widgets not as robust as ggplot2

    - does not "remember" variable names in data set
    
+ HTML can make things interactive

+ each package stores the data so can be run in java in the background

## plotly (use ggiraph)

+ allows you to see the data behind your visual

    - hover over points and see details on each data point
    
+ widget contains the data

+ larger the data, slower the response

    - not great with larger data sets unless have high computing power
    - select something in data set to cut out unnecessary data
    
+ use plotly menu and legend to zoom in and interact with graph for finer detail and filter out stuff

+ don't even really much think about doing 3-D bullshit

## dygraph (use for time series)

+ allows you to zoom in on different times with sliders

+ time series by factor level is difficult

    - must create factor time series 
    
+ good tool for working with financial data

+ versatile but sometimes difficult to work with

```{r dy1, message = FALSE, warning = FALSE}
# Convert yr_built to Date, but this is different then above when we used year()
qp1$yr_built <- as.Date(as.character(qp1$yr_built), format = "%Y")

# Summarize data
mp_ts <- qp1 %>% 
  group_by(yr_built) %>% 
  summarize(med_pr = median(price/1000))  

# Create xts data
mp_ts <- mp_ts %>% 
  xts(order.by = mp_ts$yr_built) 

# Graph median price over time
dygraph(mp_ts$med_pr) %>% 
  dyRangeSelector(dateWindow = c("1900-01-01", "2016-01-01"))
```

```{r dy2, message = FALSE, warning = FALSE}
# some easy options

# Graph median price over time
dygraph(mp_ts$med_pr, main = "Sudden decreases in value are usually larger than increases") %>% 
  dyRangeSelector(dateWindow = c("1900-01-01", "2016-01-01")) %>% 
  dyAxis("y", label = "Median price in thousands") %>% 
  dyShading(from = "1912-1-1", to = "1919-1-1", color = "#FFE6E6") %>%
  dyShading(from = "1928-1-1", to = "1933-1-1", color = "#FFE6E6") %>%
  dyShading(from = "1938-1-1", to = "1944-1-1", color = "#FFE6E6") %>%
  dyShading(from = "1970-1-1", to = "1974-1-1", color = "#CCEBD6") %>% 
  dyShading(from = "2010-1-1", to = "2016-1-1", color = "#CCEBD6")

```

```{r dy3, message = FALSE, warning = FALSE}
# time series with all quartiles

# Calculate quartiles
pr_qtile <- do.call("rbind", tapply(qp1$price/1000, qp1$yr_built, quantile)) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "time") %>% 
  mutate(time = as.Date(time)) %>% 
  rename(Min = `0%`, `2nd` = `25%`, Median = `50%`, `3rd` = `75%`, Max = `100%`)

# switch to xts format
pr_qtile_ts <- xts(x = pr_qtile[,-1], order.by = pr_qtile$time)

# Plot it
dygraph(pr_qtile_ts, main = "Max price vs quartiles shows how skewed house price data is") %>%
  dyRangeSelector(dateWindow = c("1900-01-01", "2016-01-01")) %>% 
  dyAxis("y", label = "Median price in thousands") %>% 
  dyOptions(fillGraph = TRUE)

```

```{r dy4, message = FALSE, warning = FALSE}
# time series with quartile band

# Remove min and max
pr_3tile <- pr_qtile %>% select(-c("Min", "Max"))

# switch to xts format
pr_3tile_ts <- pr_3tile[,-1] %>% 
  xts(order.by = pr_3tile$time)

# Plot
dygraph(pr_3tile_ts, main = "") %>%
  dySeries(c("2nd", "Median", "3rd")) %>% 
  dyAxis("y", label = "Median price in thousands") %>% 
  dyRangeSelector(dateWindow = c("1900-01-01", "2016-01-01")) 

```

```{r dy5, message = FALSE, warning = FALSE}
# cannot easily divide by factors

# Summarize data
ml_ts <- qp1 %>% 
  group_by(yr_built) %>% 
  summarize(med_liv = median(sqft_living), 
            med_lot = median(sqft_lot))  

# Create xts data
ml_ts <- ml_ts %>% 
  xts(order.by = ml_ts$yr_built) 

# Create factor time series
ml_ts <- cbind(ml_ts$med_liv, 
               ml_ts$med_lot)

# Graph price by waterfront over time
dygraph(ml_ts, main = "Median lot size dropped dramatically in the 1990s") %>%
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("y", label = "Square feet") %>% 
  dyRangeSelector(dateWindow = c("1900-01-01", "2016-01-01")) %>% 
  dyOptions(fillGraph = TRUE)

```

# Animations

+ help to tell the story

+ spice up some visuals

+ can be distracting if not used tastefully

+ not interactive

## Hans Rosling

+ cool video of data visualization and animation about 200 countries developing over 200 years to go from "Poor and dying" to "Rich and healthy"

+ professor Green saw this video in 2010 and decided to jump into data viz as his next focus of academia

## gganimate

+ animated visuals help deliver a story

+ works well to show motion

+ only use animations if it adds substance. don't add just for sizzle or bacon

+ use for substance, not sizzle!

### transition_time()

```{r ann1, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5}
# Build a faceted animation graph
ggplot(data = gapminder, 
       mapping = aes(x = gdpPercap/1000, y = lifeExp, 
                     size = pop, color = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  theme_bw() +

  # Here comes the gganimate specific bits
  labs(title = "Real GDP and life expectancy grow together: Year {frame_time}",
       x = "GDP per capita ($000)", y = "Life expectancy") +
  transition_time(year) +
  ease_aes("linear")

```

### transition_reveal()

```{r ann2, message = FALSE, warning = FALSE}
# Graph median price by year built
p <- qp1 %>% 
  select(yr_built, waterfront, price) %>% 
  group_by(yr_built, waterfront) %>% 
  summarize(med_price = median(price/1000)) %>% 
  ggplot(aes(x = yr_built, y = med_price, frame = yr_built)) + 
    geom_line(aes(color = as.factor(waterfront))) + 
    scale_color_manual(name = "Waterfront", labels = c("No", "Yes"), values = c("navy", "blue"),
                       guide = guide_legend(reverse=TRUE)) +
    theme_classic() +
    theme(axis.title.x = element_blank(),
          legend.position="bottom") +
  labs(title = "Waterfront median price is much more variable than non-waterfront",
       x = "", y = "Median price ($000)") +  
  transition_reveal(yr_built) 

# Run as a GIF animation and control speed
animate(p, nframes = 300)

```


### transition_state()

```{r ann3, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 5}
# Animate a bar graph
qp1 %>% 
  select(price, bedrooms, waterfront) %>% 
  group_by(bedrooms, waterfront) %>%
  filter(bedrooms < 7 & bedrooms > 0) %>% 
  summarise(med_price = median(price/1000)) %>% 
  ggplot(aes(x = bedrooms, y = med_price, fill = as.factor(waterfront))) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  scale_fill_manual(name = "Waterfront", labels = c("No", "Yes"), values = c("lightblue", "blue"),
                    guide = guide_legend(reverse=TRUE)) +
  labs(title = "Median price grows more per bedroom for waterfront homes",
       x = "Number of bedrooms", y = "Median price ($000)") +
  transition_states(bedrooms, wrap = FALSE) +
  shadow_mark()

```


# shiny

+ interactive visual page that can be printed

+ user is changing variables for the perspectives on the data

+ HTML widgets only allow for viewing static data on the page

    - doesn't allow you to change the variables or change the data
    - only zooms in and out
    - only impacts "hover height" of the data
    
+ best practices

    1. keep it clear
        - keep it simple
        - choose visuals carefully
            - bar graphs: categoricla data
            - line graphs: time series and variables that show motion
            - scatter plot: show relationships between variables
            - maps: show spatial relationships
        
    2. make it approachable
        - arrange interactions with simplicity in mind
        - make the audience the focus, rather than the analysis
        - focus on what the audience needs to know, not the analytics
        - focus on the finding, not the method
        
## shiny interactive visuals

1. example 1: 

    - Each dashboard element is driven by 1 menu
    - shareable on social media
    - embed code into dashboard in the background
    - informational dashboard. only allows changes to 3 key variables independently, but not together
        - does not allow for as deep of a dive
    - master chunk will allow user to dive deeper
    
2. example 2: 

    - statistical analysis of sampling distributions of Central Limit Theorem 
    - shiny structure: 
    - 3 elements of shiny dashboards:
        1. inputPanel() - create space on page and gather user input
        2. reactive[()] - filter data with user input
        3. renderPlot({}) - plot data determined by user input
    - using stuff to dive deeper into the data
        
    
3. example 3: 

    - master chunk drives updates to all elements
    
    

